---
title: "Convert Master Grid CSV back to Geotiff"
author: "Lyndon Estes"
date: "7/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Converting the CSV file of the master grid back to a geotiff is fairly memory intensive, and doesn't seem to have a ready solution with gdal (I tried gdal_grid but it didn't work).  

It also couldn't be done on a local machine with <32 GB RAM, so the solution was to spin up an AWS instance with R and run it on that.

## AWS set up

Following advice on AWS re. R on instance.

```
sudo apt-get update
sudo apt-get upgrade

# get the right version of R
# from here: https://askubuntu.com/questions/909689/upgrading-r-version-3-3-in-ubuntu-16-04
# because current Ubuntu version to old for rgdal
sudo add-apt-repository ppa:marutter/rrutter
sudo apt update
sudo apt full-upgrade

# gdal
# from https://gis.stackexchange.com/questions/193814/installing-gdal2-1/193828#193828
sudo add-apt-repository -y ppa:ubuntugis/ubuntugis-unstable
sudo apt update 
sudo apt install gdal-bin python-gdal python3-gdal

# also needed this to get rgdal to install
sudo apt-get install libgdal1-dev libproj-dev gdal-bin proj-bin

# then R packages
sudo R -e "install.packages('rgdal', repos='http://cran.r-project.org')"
sudo R -e "install.packages('devtools', repos='http://cran.rstudio.com/')"
sudo R -e "devtools::install_github('ldemaz/dtraster')"

# then create directories with scripts and ship csv to instance
mkdir data
mkdir R
```

scp commands to send data to instance
```
cd ~/Dropbox/projects/activelearning/mapperAL/spatial/data/processed
scp -i ~/Desktop/lde2.pem africa_master_grid_newf.csv ubuntu@18.232.105.154:data/

```

## R code on instance

This ran, but `dt_to_raster` failed to re-grid accurately, so there was offset between old and new grids. 

```{r, eval = FALSE}
library(dtraster)
setwd("/home/ubuntu/data")
mgrid <- fread("africa_master_grid_newf.csv")
# pth <- "spatial/data/processed/africa_master_grid_newf.csv"
# mgrid <- fread(pth)

# adjust database
mgrid[, fwts := NULL]
mgrid[, cntry := substr(name, 1, 2)]
mgrid[, cntry := as.integer(as.factor(cntry)))]
mgrid[, cntry_id := substr(name, 3, 9)]
mgrid[, cntry_id := as.integer(cntry_id)]

str(mgrid[1:10, ])
# # initial writeout of ID, for test
gcsstr <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
# a <- dt_to_raster(mgrid[, .(x, y, ID)], CRSobj = gcsstr, 
#                   filename = "mgrid_id.tif")
# rm(a)
# gc()

# # write out country code and country ID
# b <- dt_to_raster(mgrid[, .(x, y, cntry, cntry_id)], CRSobj = gcsstr, 
#                   filename = "mgrid_cid.tif")
# rm(b)
# gc()

# write out seasons
#nms <- "cntry_id"#c("cntry")#, "cntry_id", 
nms <- c("dry_start", "dry_end", "wet_start", "wet_end")
for(i in nms) {
  print(paste("rasterizing", i))
  b <- dt_to_raster(mgrid[, c("x", "y", i), with = FALSE], CRSobj = gcsstr,
                    filename = paste0("mgrid_", i, ".tif"))
  rm(b)
  gc()
}

```

### Bring back results
```
scp -i ~/Desktop/lde2.pem ubuntu@18.232.105.154:data/mgrid* .
```

### Combine into multiband
```{r, eval = FALSE}
library(gdalUtils)
p_int <- "spatial/data/interim"
p_proc <- "spatial/data/processed"

nms <- c("id", "cntry", "cntry_id", "dry_start", "dry_end", "wet_start",
         "wet_end")
nmsv <- unname(unlist(sapply(nms, function(x) {
  dir(p_int, pattern = paste0("mgrid_", x, ".tif"), full.names = TRUE)
})))
otypes <- c("UInt32", "Byte", "UInt32", "Byte", "Byte", "Byte", "Byte")

# change data types
for(i in 1:length(nmsv)) {
  onm <- paste0(p_int, "/", paste0("mgrid_", nms[i] , "_t.tif"))
  print(onm)
  gdal_translate(src_dataset = nmsv[i], dst_dataset = onm, overwrite = TRUE, 
                 ot = otypes[i], a_nodata = 0)
}

nmsv <- unname(unlist(sapply(nms, function(x) {
  dir(p_int, pattern = paste0("mgrid_", x, "_t.tif"), full.names = TRUE)
})))
vnm <- paste0(p_int, "/", paste0("mgrid_all.vrt"))
gdalbuildvrt(gdalfile = nmsv, separate = TRUE, overwrite = TRUE,
               output.vrt = vnm)
tnm <- paste0(p_proc, "/africa_master_grid_allf.tif")
gdal_translate(src_dataset = vnm, dst_dataset = tnm, overwrite = TRUE, 
               ot = "UInt32")

```

`gdal_translate` led to a gigantic output geotiff, and there was offset, so a different solution was needed. 

## Fix

Since the approach with `dt_to_raster` didn't work, going back to approach based on raster package. 

```{r, eval = FALSE, echo = FALSE}
# Testing that it works first
library(sf)
mgrid_new <- brick(tnm)
# mgrid_old <- brick("spatial/data/processed/archive/africa_master_brick_all.tif")
e <- st_bbox(mgrid_new)
# st_bbox(mgrid_new) == st_bbox(mgrid_old)
# paste(diff(e[c(1, 3)]) / 0.005)
# paste(diff(e[c(2, 4)]) / 0.005)

# mgrid[, diff(range(x)) / 0.005]
# mgrid[, diff(range(y)) / 0.005]

s <- mgrid[sample(1:.N, 100), ]

cid <- cellFromXY(mgrid_new[[1]], s[, .(x, y)])
# eext <- extract(mgrid_new, s[, .(x, y)])
eext <- mgrid_new[cid]
cbind(s$ID, eext[, 1])
s$ID - eext[, 1]
class(s$ID)

# test against old grid
s2 <- mgrid[sample(1:.N, 100), ]
cid2 <- cellFromXY(mgrid_old[[1]], s2[, .(x, y)])
eext2 <- mgrid_old[cid2]
cbind(s2$ID, eext2[, 1])
all(s2$ID == eext2[, 1])

# does it write out accurately
cid <- cellFromXY(mgrid_old[[1]], mgrid[, .(x, y)])
r <- mgrid_old[[1]]
r[cid] <- mgrid[, ID]
writeRaster(r, filename = paste0("spatial/data/interim/mgrid_", "id", "f.tif"), 
            datatype = "INT4U")
r - mgrid_old[[1]]
```

```{r, echo = FALSE, eval = FALSE}
# clean-up old files
p_int <- "~/Dropbox/projects/activelearning/mapperAL/spatial/data/interim"
file.remove(dir(p_int, pattern = "_t.tif", full.names = TRUE))
file.remove(dir(p_int, pattern = "all.vrt", full.names = TRUE))
# moved manually dt_to_raster processed files to archive folder in interim
```

Run this on the m5.4xlarge instance. Code:
```{r, eval = FALSE}
setwd("/home/ubuntu/data")
library(dtraster)
mgrid <- fread("africa_master_grid_newf.csv")

# adjust database
mgrid[, fwts := NULL]
mgrid[, cntry := substr(name, 1, 2)]
mgrid[, cntry := as.integer(as.factor(cntry))]
mgrid[, cntry_id := substr(name, 3, 9)]
mgrid[, cntry_id := as.integer(cntry_id)]

# extent of our grid
e <- c(-17.541, -34.845, 51.419, 37.540)
names(e) <- c("xmin", "ymin", "xmax", "ymax")
r <- raster(extent(e[c(1, 3, 2, 4)]))  # fresh new dummy raster
res(r) <- 0.005

cid <- cellFromXY(r, mgrid[, .(x, y)])  # cell number

# mgrid[cntry == 1]
nms <- c("ID", "cntry", "cntry_id", "dry_start", "dry_end", "wet_start",
         "wet_end")
for(i in nms) {
  f <- paste0("mgrid_", i, "f.tif")
  print(f)
  r[cid] <- mgrid[[i]]
  writeRaster(r, filename = f, datatype = "INT4U")
}

# write a single brick
fnms <- paste0("mgrid_", nms, "f.tif")
s <- stack(lapply(fnms, raster))
br <- brick(s, filename = "africa_master_brick_new.tif", overwrite = TRUE, 
            datatype = "INT4U") # change datatype
```

Bring it back (from terminal)
```
cd ~/Dropbox/projects/activelearning/mapperAL/spatial/data/interim
scp -i ~/Desktop/lde2.pem ubuntu@34.231.21.32:data/mgrid*f.tif .
cd ~/Dropbox/projects/activelearning/mapperAL/spatial/data/processed
scp -i ~/Desktop/lde2.pem ubuntu@34.231.21.32:data/africa_master_brick_new.tif africa_master_grid_newf.tif 
```

## Checks

To see if everything lines up. In QGIS, ID didn't seem to initially line up (a look back shows it did). Running here a number of checks against the csv file, as well as the existing master_grid database. 

```{r, eval = FALSE}
library(raster)
b <- brick("spatial/data/processed/africa_master_grid_newf.tif")
s <- mgrid[sample(1:.N, 100), ]
cid <- cellFromXY(b[[1]], s[, .(x, y)])
bvals <- b[cid]

s$ID == bvals[, 1]
s$cntry_id == bvals[, 3]
s$wet_start == bvals[, 6]
s$dry_end == bvals[, 5]
s$cntry == bvals[, 2]
as.integer(substr(s$name, 3, 9))  == bvals[, 3]
```

Against current master grid and existing Q, F, and Meridia sites 
```{r, eval = FALSE}
library(rmapaccuracy)
library(sf)
coninfo <- mapper_connect(user = pgupw$user, password = pgupw$password,
                          db.tester.name = "lestes", 
                          host = "crowdmapper.org")
qnms <- (tbl(coninfo$con, "kml_data") %>% filter(kml_type == "Q") %>% 
           collect())$name
q_in_mgrid <- (tbl(coninfo$con, "master_grid") %>% 
                 filter(name %in% qnms) %>% collect())

# as.integer(substr(q_in_mgrid$name, 3, 9)) 
qmgrid <- mgrid[name %in% qnms]
qmgrid <- qmgrid[order(name)]
q_in_mgrid <- as.data.table(q_in_mgrid)
q_in_mgrid <- q_in_mgrid[order(name)]
all(qmgrid$x == q_in_mgrid$x)
all(qmgrid$y == q_in_mgrid$y)
qmgrid$name == q_in_mgrid$name
qmgrid$ID == q_in_mgrid$id

# F Names
fnms <- (tbl(coninfo$con, "kml_data") %>% filter(kml_type == "F") %>% 
           collect())$name
f_in_mgrid <- (tbl(coninfo$con, "master_grid") %>% 
                 filter(name %in% fnms) %>% collect())

# as.integer(substr(f_in_mgrid$name, 3, 9)) 
fmgrid <- mgrid[name %in% fnms]
fmgrid <- fmgrid[order(name)]
f_in_mgrid <- as.data.table(f_in_mgrid)
f_in_mgrid <- f_in_mgrid[order(name)]
all(round(fmgrid$x, 4) == round(f_in_mgrid$x, 4))
all(round(fmgrid$y, 4) == round(f_in_mgrid$y, 4))
fmgrid$name == f_in_mgrid$name
fmgrid$ID == f_in_mgrid$id

# check against Ghana selected grids
gh <- read_sf("spatial/data/processed/ghanasites.kml")
mgridgh <- mgrid[name %in% gh$Name]

# pols versus mgrid
towns <- unique((gh %>% select(town))$town)
par(mfrow = c(2, 3), mar = rep(0, 4))
for(i in towns) {
  gcells <- st_geometry(gh %>% filter(town == i))
  plot(gcells)
  plot(as(extent(as_Spatial(gcells)), "SpatialPolygons"), add = TRUE)
  mtext(side = 3, text = i, line = -1)
  towngr <- mgridgh[name %in% (gh %>% filter(town == i))$Name, ]
  towngr[, points(x, y)]
  gpol <- (gh %>% filter(town == i))[sample(1:20, 1), ]
  plot(st_geometry(gpol), add = TRUE, col = "blue")
  towngr[name %in% gpol$Name, points(x, y, pch = 20, col = "red")]
}

# Meridia names from current master_grid
gh_in_mgrid <- (tbl(coninfo$con, "master_grid") %>% 
                  filter(name %in% gh$Name) %>% collect())
gh_in_mgrid <- gh_in_mgrid %>% arrange(name)
mgridgh <- mgridgh[order(name)]

all(mgridgh[, name] == gh_in_mgrid$name)  # names match
all(mgridgh[, ID] == gh_in_mgrid$id)  # ids match

```

Last one is to check the IDs of the coastal zones relative to the new IDs
```{r, eval = FALSE}
br_in_mgrid <- (tbl(coninfo$con, "master_grid") %>% 
                  filter(grepl("BR", name)) %>% collect())
mgrid_coastal <- mgrid[ID %in% br_in_mgrid$id]
s <- mgrid_coastal[sample(1:.N, 10), .(ID, x, y)][order(ID)]
s2 <- br_in_mgrid %>% filter(id %in% s$ID) %>% select(id, x, y) %>% arrange(id)
all(s$x == s2$x)
all(s$y == s2$y)

# last confirm that these are the same cells in the new raster
b <- brick("spatial/data/processed/africa_master_grid_newf.tif")
cid <- cellFromXY(b[[1]], mgrid_coastal[, .(x, y)])
bvals <- b[cid]
all(mgrid_coastal$ID == bvals[, 1])  # yes
```

Okay, so looks like everything lines up. 

Clean-up files
```{r, eval = FALSE}
file.copy("spatial/data/interim/africa_master_brick.tif", 
          "spatial/data/interim/archive/africa_master_brick.tif")
file.remove("spatial/data/interim/africa_master_brick.tif")
file.copy("spatial/data/processed/africa_master_grid_allf.tif", 
          "spatial/data/processed/archive/africa_master_grid_allf.tif")
file.remove("spatial/data/processed/africa_master_grid_allf.tif*")
file.remove("spatial/data/processed/mgrid_id.tif")
file.copy("spatial/data/processed/africa_master_grid_new.csv", 
          "spatial/data/processed/archive/africa_master_grid_new.csv")
file.remove("spatial/data/processed/africa_master_grid_new.csv") 
```


