---
title: "Make New Master Grid"
author: "Su Ye & Lei Song"
date: "2/16/2018"
output: 
  html_document:
    theme: united
    number_sections: true 
    highlight: tango
---

```{r setup, include=FALSE, echo = TRUE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

## Pre-work

### Install the essential packages and load them into environment

1. `raster`, `rgdal`, `rgeos`, and `gdalUtils` are used for geographical data.

2. `lmisc` is a package to manage the working folder and the path more easily.

3. `dtraster` is a package for 'as.data.table' function.

4. Package `rmapaccuracy` is a package built in the project, need to be tailored accordingly. A good way to install it is to download the github repo, change the code and install locally.

4. Package `lmisc`, `dtraster` are Github packages. Need to be installed by `install_github()` in package `devtools`.

```{r packages, message=FALSE, warning=FALSE, eval = TRUE}
library(raster) # packages for geographical data
library(rgdal)
library(rgeos)
library(gdalUtils)
library(lmisc) # package for handling the working directory and path
require(dtraster)
library(dplyr) # package for big database
library(dbplyr)
library(RPostgreSQL)
library(rmapaccuracy) # package for KML generation
library(yaml)

params  <-yaml.load_file(file.path(Sys.getenv('PYTHONPATH'),'config.yaml'))
```

### Check the version of packages and R

It is necessary to update the GDAL to the recent version. Different versions of GDAL might use different methods to process the data. Slightly difference could be a disaster for geographical data. 

For example, an earlier version of gdal caused a raster reprojection to have slightly different results compared to the later version.

```{r checkVersion, message=FALSE, warning=FALSE, eval = TRUE}
sessionInfo() # Check the system information
gdal_setInstallation(rescan = FALSE) # Check the GDAL
## check GDAL version
message("GDAL version is: ")
getOption("gdalUtils_gdalPath")[[1]]$version
```
**NOTE:** GDAL version should be 2.2.2 or higher.

### Set the working directory

All the functions here come from package `lmisc`. This package helps us to manage the working folder and the path easily.

```{r wd, message=FALSE, warning=FALSE, eval = TRUE}
## Set the working directory
p_extdat <- "spatial/data/external/"
p_intdat <- "spatial/data/interim/"
p_procdat <- "spatial/data/processed/"
```

## Load Africa boundary line and buffer by 500 m (1 pixel)

### Fetch Africa with GCS projection

```{r load, message=FALSE, warning=FALSE, eval = FALSE}
af <- readOGR(dsn = paste0(p_extdat, "africa_countries_alb.sqlite"), 
              layer = "africa_countries_alb", verbose = FALSE)

af_cnt <- readOGR(dsn = paste0(p_extdat, "africa_noisl_gcs.sqlite"), 
                  layer = "africa_noisl_gcs", verbose = FALSE)
wgs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
af_cnt@proj4string <- CRS(wgs)
```

### Buffer 1 km for the boundary

Before we do buffer, all layers' projection should be Albers, since the function `gBuffer` only works for projected coordination. After we finish the buffer, change the projection back to GCS.

```{r buffer, message=FALSE, warning=FALSE, eval = FALSE}
af_cnt_alb <- spTransform(af_cnt, af@proj4string)  # transform to Albers
af_alb <- gUnaryUnion(af_cnt_alb)  # dissolve countries
af_alb <- SpatialPolygonsDataFrame(af_alb, data = data.frame("ID" = 1), 
                                   match.ID = FALSE)
afbuf <- gBuffer(af_alb, width = 500)  # buffer by 500 m
afbuf <- SpatialPolygonsDataFrame(afbuf, 
                                  data = data.frame("ID" = 1), 
                                  match.ID = FALSE)

## Transform back to GCS projection
af <- spTransform(af_alb, CRS(wgs))
afbuf <- spTransform(afbuf, CRS(wgs))
rm(af_cnt_alb, af_alb)
```

### Save out the files

```{r saveORG, message=FALSE, warning=FALSE, eval = FALSE}
writeOGR(af_cnt, 
         dsn = fp(p_intdat, "africa_countries_gcs.sqlite"), 
         layer = "africa_countries_gcs", 
         driver = "SQLite", dataset_options = c("SPATIALITE = yes"),
         overwrite_layer=TRUE)
writeOGR(afbuf, 
         dsn = fp(p_intdat, "africa_gcs_500m_buff.sqlite"), layer = "afbuf",
         driver = "SQLite", dataset_options = c("SPATIALITE = yes"),
         overwrite_layer=TRUE)
```

## Grid the buffered shape

Get the master grid with 0.005 degree * 0.005 degree. 'buffrast.tif' is a binary image for Africa area, while "africa_master_grid.tif" is a image for Africa area. Value of each pixel is the pixel's unique ID.

**NOTE:** When we use `writeRaster`, be careful of datatype. Define the datatype to a proper one. Because `R` won't give you an error for these, instead it will mess your stuff up.

```{r grid, message=FALSE, warning=FALSE, eval = FALSE}
gdal_rasterize(src_datasource = paste0(p_intdat, "africa_gcs_500m_buff.sqlite"),
               ot = "Int16", tr = c(0.005, 0.005), at = TRUE, a_nodata = 0,
               dst_filename = paste0(p_intdat, "buffrast.tif"),
               te = c(-17.541, -34.846, 51.420, 37.540),
               burn = 1, l = "afbuf", of = "GTiff") 
afgrid <- raster(paste0(p_intdat, "buffrast.tif"))

cells <- Which(!is.na(afgrid), cells = TRUE)
ids <- 1:length(cells)
afgrid[cells] <- ids # make the values of each grid is its index
afgrid <- writeRaster(afgrid, datatype = "INT4U", # datatype should be INT4U
                      filename = fp(p_intdat, "africa_master_grid.tif"), 
                      overwrite = TRUE)
```

## Process cropland data for sample weights

First read the cropland percentage image got from Geowiki. Then do classification to the raw image. Of course, we use the country boundary to tailor the final image.

```{r fwts, message=FALSE, warning=FALSE, eval = FALSE}
f <- paste0(p_extdat, "cropland/Hybrid_10042015v9.img")
fo <- paste0(p_intdat, "af_cropland.tif")
afgridbuf <- raster(paste0(p_intdat, "buffrast.tif"))

gdalwarp(srcfile =  f, t_srs = projection(afgrid), dstfile = fo,
         r = "bilinear", ot = "Float32", te = bbox(afgrid)[1:4],
         dstnodata = -32768, tr = res(afgrid), of = "GTiff",
         verbose = TRUE, overwrite = TRUE) # get subset of cropland for africa

cropland <- raster(fo)
cropland_p <- cropland / 100 * afgridbuf # calculate the percentage

recl <- cbind(seq(0, 1, 0.1)[-11], seq(0, 1, 0.1)[-1], 1:10)
f <- paste0(p_intdat, "af_cropland_prob_cl.tif")
cropland_pcl <- reclassify(cropland_p, rcl = recl, include.lowest = TRUE,
                           filename = f, overwrite=TRUE) # do relcassification to cropland percentage data

```

## Create the raster brick

Before we save out the raster brick, need to make sure the structure of image `afgrid` and `cntr` are identical so that they can match each other. In this case, the buffer is an issue. So some preprocess is essential.

1. Rasterize the country boundaries.

2. Merge the result and the binary image for Africa area 'buffrast.tif' to make a image of country index with buffer. The value of each pixel is the ID of the country this pixel belonging to.

3. Save the master grid raster and the image of country index with buffer as a brick.

**NOTE:** When we use `brick`, be careful of datatype. Define the datatype to a proper one.

```{r brick, message=FALSE, warning=FALSE, eval = FALSE}
## rasterize the countries
af_f <- paste0(p_intdat, "africa_countries_gcs.sqlite")
af_of <- paste0(p_intdat, "africa_countries_gcs.tif")
afgridbuf <- raster(paste0(p_intdat, "buffrast.tif"))
gdal_rasterize(src_datasource = af_f, dst_filename = af_of, at = TRUE,
               ot = "Int16", a = "objectid", l = "africa_countries_gcs",
               tr = res(afgrid), verbose = FALSE, te = bbox(afgrid)[1:4],
               a_nodata = -32768, of = "GTiff")

## merge the buffer back to the afgrid
cntr <- raster(af_of)
cntr <- merge(cntr, afgridbuf, incomparables = 1)

cntrm  <- cntr >= 1
afgridr <- afgrid * cntrm
cropland_pcl <- cropland_pcl * cntrm

## create complete master
st <- stack(list(afgridr, cropland_pcl, cntr))
br <- brick(st, filename = paste0(p_intdat, "africa_master_brick.tif"), 
            overwrite = TRUE, datatype = "INT4U") # change datatype
```

## Create complete master grid textfile

```{r mgrid, message=FALSE, warning=FALSE, eval = FALSE}
## give buffer surrounding the boundary a 'iso2' - BR
attr <- af_cnt@data[, c("objectid", "iso2")]
attr$iso2 <- as.character(attr$iso2)
attr <- rbind(attr, c(1, "BR"))
attr$iso2 <- as.factor(attr$iso2)
attr$objectid <- as.numeric(attr$objectid)
af_cnt_dt <- data.table(attr, key = "objectid")
setnames(af_cnt_dt, "objectid", "cnt_code")

## read brick into a table and clean it
br_dt <- as.data.table(br, xy = TRUE, progress = "text")
setnames(br_dt, old = colnames(br_dt)[3:5], 
         new = c("ID", "fwts", "cnt_code"))
br_dt <- br_dt[!is.na(ID)]

## create unique ID number for each cell in each country
for(i in unique(br_dt$cnt_code)) {
        ind <- which(br_dt$cnt_code == i)
        br_dt[ind, ind := order(ind)]
}

setkey(br_dt, cnt_code)
mgrid <- merge(br_dt, af_cnt_dt, by = "cnt_code")
ndig <- nchar(max(mgrid$ind))
mgrid[, name := sprintf(paste0("%s%0", ndig, "i"), iso2, ind)]
setcolorder(mgrid, c("ID", "x", "y", "name", "fwts", "cnt_code", "iso2",
                     "ind"))
mgrid[, (c("cnt_code", "iso2", "ind")) := NULL]

## change data types
names_int <- c("ID", "fwts")
for(col in names_int) set(mgrid, j = col, value = as.integer(mgrid[[col]]))
mgrid[, avail := rep("T", nrow(mgrid))]

## change decimal
mgrid[, x := round(x, 4)]
mgrid[, y := round(y, 4)]

## write out the table into a csv file
fnm <- paste0(p_procdat, "africa_master_grid_new.csv")
fwrite(mgrid, file = fnm, col.names = TRUE, row.names = FALSE, verbose = FALSE)
```

### Clean-up interim files

Leave behind `africa_master_brick.tif` in interim folder. 
```{r, eval = FALSE}
frm <- paste0(p_intdat, c("af_cropland.tif", "af_cropland_prob_cl.tif", 
                          "africa_countries_gcs.sqlite", 
                          "africa_countries_gcs.tif", 
                          "africa_master_grid.tif", 
                          "africa_gcs_500m_buff.sqlite", 
                          "buffrast.tif"))
file.remove(frm)
```


## Write DT into postgis

The database is already built in our server. The name is 'AfricaSandbox'. Use `dbplyr` package to connect to the database by defining the correct `host`, `port`, `user`, and `password`. 

```{r postgis, message=FALSE, warning=FALSE, eval = FALSE}
## create the database
# system("createdb AfricaSandbox")

## create user
# system(paste("createuser -s -l -d -r ", params$labeller$db_username))

## connect to the database
con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(),
                      dbname = "AfricaSandbox",   
                      user = params$labeller$db_username,
                      password = params$labeller$db_password)

## create a table in database
sql <- paste("CREATE TABLE master_grid_new",
             "(gid serial PRIMARY KEY,",
             "ID integer, x double precision, y double precision,",
             "name char(9), fwts integer, avail char(1));")
dbSendQuery(con, sql)

## copy csv file into the table
sql <- paste0("\\COPY master_grid_new(ID, x, y, name, fwts, avail) ",
             "FROM '", fnm, "' WITH DELIMITER ',' CSV HEADER;")
dbSendQuery(con, sql)

# improve the efficiency of querying
dbSendQuery(con, "CREATE INDEX name_gix_new ON master_grid_new (name);")
dbSendQuery(con, "VACUUM ANALYZE master_grid_new;")
dbSendQuery(con, "CLUSTER master_grid_new USING name_gix_new;")

## close the database
DBI::dbDisconnect(con)
```

## Check the final master grid

### Querying efficiency

```{r efficiency, message=FALSE, warning=FALSE, eval = TRUE}
con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(),
                      dbname = "AfricaSandbox",   
                      user = params$labeller$db_username,
                      password = params$labeller$db_password)

# ids <- sample(mgrid$ID, 20 ,replace=FALSE) # get random ID from mgrid to test

ids <- 94649043

ptm <- proc.time()
pixels <- tbl(con, "master_grid_new") %>%
        filter(id == ids) %>%
        collect()
proc.time() - ptm

DBI::dbDisconnect(con)
```

### Select different clusters across the whole Africa

```{r clusters, message=FALSE, warning=FALSE, eval = TRUE, fig.width=9.5, fig.height=9.5}
library(leaflet)
africa <- readOGR(fp(p_odat, "africa_countries_gcs.sqlite"), verbose = FALSE)
con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(),
                      dbname = "AfricaSandbox",   
                      user = params$labeller$db_username,
                      password = params$labeller$db_password)

## set parameters
wgs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
diam <- 0.005/2

## Put all functions together
kml_generate <- function(pixels){
        xy_tabs <- data.table(pixels)
        
        ## Step 1. Convert point data to proper projections
        gpnts <- sapply(1:nrow(xy_tabs), function(i) {  # i <- 1 
                xy <- data.frame(xy_tabs[i, ])
                coordinates(xy) <- ~x + y # xy[1:2]
                # xy <- xy[, 3:5]
                crs(xy) <- CRS(wgs)
                xy
        })
        
        ## Step 2. Convert points to polygons
        gpols <- sapply(1:nrow(xy_tabs), function(i) { 
                xy_trans <- gpnts[[i]] 
                proj4 <- proj4string(xy_trans)
                ptdata <- data.frame(xy_trans)
                gpol <- point_to_gridpoly(xy = ptdata, w = diam, 
                                          CRSobj = CRS(proj4))
                gpol@data$id <- i
                gpol <- spChFIDs(gpol, as.character(gpol@data$id))
        })
        
        ## Step 3. Transform to geographic coordinates and write out to kmls
        gpols_gcs <- sapply(1:length(gpols), function(i) {
                gpol_gcs <- gpols[[i]]
                kml_name <- paste0("/home/lsong/mappingafrica/data/kmls/",
                                   gpol_gcs$name, ".kml")
                rgdal::writeOGR(gpol_gcs, dsn = kml_name, layer = gpol_gcs$name,                                 driver = "KML", overwrite = TRUE)
                gpol_gcs
        })
        gpols_gcs <- do.call(rbind, gpols_gcs)
}

## north-west
pixels <- tbl(con, "master_grid_new") %>%
        filter((id == 26006600) | (id == 26006601) | (id == 26017292) | 
                 (id == 26017293)) %>% collect()

pixels$name
nw <- kml_generate(pixels)

## north-center
pixels <- tbl(con, "master_grid_new") %>%
        filter((id == 11961260) | (id == 11961261) | (id == 11970741) | 
                 (id == 11970742)) %>% collect()

pixels$name
nc <- kml_generate(pixels)

## north-east
pixels <- tbl(con, "master_grid_new") %>%
        filter((id == 7688127) | (id == 7688128) | (id == 7697081) | 
                 (id == 7697082)) %>% collect()

pixels$name
ne <- kml_generate(pixels)

## center-west
pixels <- tbl(con, "master_grid_new") %>%
        filter((id == 69102813) | (id == 69102814) | (id == 69109497) | 
                 (id == 69109498)) %>% collect()

pixels$name
cw <- kml_generate(pixels)

## center-center
pixels <- tbl(con, "master_grid_new") %>%
        filter((id == 62394861) | (id == 62394862) | (id == 62403237) | 
                 (id == 62403238)) %>% collect()

pixels$name
cc <- kml_generate(pixels)

## center-east
pixels <- tbl(con, "master_grid_new") %>%
        filter((id == 60618428) | (id == 60618429) | (id == 60628923) | 
                 (id == 60628924)) %>% collect()

pixels$name
ce <- kml_generate(pixels)

## south-west
pixels <- tbl(con, "master_grid_new") %>%
        filter((id == 100287476) | (id == 100287477) | (id == 100290355) | 
                 (id == 100290356)) %>% collect()

pixels$name
sw <- kml_generate(pixels)

## south-center
pixels <- tbl(con, "master_grid_new") %>%
        filter((id == 99875112) | (id == 99875113) | (id == 99878249) | 
                 (id == 99878250)) %>% collect()

pixels$name
sc <- kml_generate(pixels)

## south-east
pixels <- tbl(con, "master_grid_new") %>%
        filter((id == 98236189) | (id == 98236190) | (id == 98239788) | 
                 (id == 98239789)) %>% collect()

pixels$name
se <- kml_generate(pixels)

fc <- c("transparent", "red", "blue")
m <- leaflet() %>% 
        addTiles() %>% 
        addPolygons(data = africa, weight = 2, fillColor = fc[1], 
                    color = fc[2]) %>%
        addPolygons(data = nw, weight = 1, fillColor = fc[1], color = fc[3]) %>%
        addPolygons(data = nc, weight = 1, fillColor = fc[1], color = fc[3]) %>%
        addPolygons(data = ne, weight = 1, fillColor = fc[1], color = fc[3]) %>%
        addPolygons(data = cw, weight = 1, fillColor = fc[1], color = fc[3]) %>%
        addPolygons(data = cc, weight = 1, fillColor = fc[1], color = fc[3]) %>%
        addPolygons(data = ce, weight = 1, fillColor = fc[1], color = fc[3]) %>%
        addPolygons(data = sw, weight = 1, fillColor = fc[1], color = fc[3]) %>%
        addPolygons(data = sc, weight = 1, fillColor = fc[1], color = fc[3]) %>%
        addPolygons(data = se, weight = 1, fillColor = fc[1], color = fc[3])
m

DBI::dbDisconnect(con)
```
