---
title: "Replace existing Q/I sites"
author: "Lyndon Estes & Lei Song"
date: "3/6/2018"
output: 
  html_document:
    theme: united
    number_sections: true 
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval = FALSE)
```

## Pre-work

Load the essential packages and define the constants.

```{r prework}
library(rgdal)
library(rgeos)
library(sf)
library(dplyr)
library(glue)
library(data.table)
library(raster)
library(gdalUtils)
library(yaml)
params <- yaml.load_file(file.path(Sys.getenv('PYTHONPATH'),'config.yaml'))

p_kmls <- "kmls_static"
p_grid <- "spatial/data/interim/"
p_dat <- "spatial/data/external/"
p_old <- "external/production_062016/data/07112017/"
p_oldg <- "external/ext_data/"
p_newg <- "external/ext_data/"
alb <- paste("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 +y_0=0",
             "+ellps=WGS84 +datum=WGS84 +units=m +no_defs") 
epsg <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

## Custom Albers read-in from mapper database, for Zambia kmls from old grid
load(file.path(p_dat, "prjstr.rda"))
```

## Get centriods

Read in KMLs of static file system and get the centroids for them.

```{r getCentriods}
## Read-in centroids
kmlcentroids <- lapply(dir(p_kmls, full.names = TRUE), function(x) {
        kml <- read_sf(x)
        pt <- st_transform(st_centroid(st_transform(kml, alb)), st_crs(kml))
        pt <- st_bbox(pt)[1:2]
        data.frame("name" = kml$Name, "x" = unname(pt[1]), "y" = unname(pt[2]), 
             stringsAsFactors = FALSE)
})
ptmat <- do.call(rbind, kmlcentroids)

## Grab the old kml sites
oldkmls <- fread(file.path(p_old, "kmls.csv"))
zmsites <- oldkmls[kml_type == "Q" & like(name, "ZM")]
```

## Align centriolds

Extract cell IDs from new master grid brick, as well as the left behind Zambia ones from the old master brick.

```{r alignCentriods}
mgrid <- raster(file.path(p_grid, "africa_master_brick.tif"))  # new grid tif file
mgrido <- fread(file.path(p_oldg, "africa_master_grid.csv"))  # old grid csv file

## First, intersect points with the old master brick to get handful of Zambia 
# Q sites
zmq <- mgrido[name %in% zmsites$name]
zmq <- zmq[, list(x, y, name, ID, zone)]
setnames(zmq, "ID", "id")
coordinates(zmq) <- ~x + y
proj4string(zmq) <- alb

zmq <- spTransform(zmq, prjstr[unique(zmq$zone)])
zmq <- spTransform(zmq, crs(mgrid))
zmq@data <- zmq@data[, c("name"), drop = FALSE]

## Second, get the kmls
coordinates(ptmat) <- ~x + y
proj4string(ptmat) <- crs(mgrid)
ptmat <- rbind(ptmat, zmq)

ptbufs <- gBuffer(ptmat, width = 0.011, byid = TRUE)

# Cells in master grid overlaped with ptbufs
vals <- extract(mgrid, ptbufs, cellnumbers = TRUE)
valsdf <- do.call(rbind, vals)

# Clean the duplicated cells
ids <- unique(valsdf[, 2])
```

## Get the names of overlapped cells

Based on cell IDs to get their names, then save the names into database.

```{r getname}
min_ids <- min(ids)

# Get the names of old kmls
mgridn <- fread(file.path(p_newg, "africa_master_grid_new.csv")) # new grid csv file
mgridn_subset <- mgridn[ID >= min_ids - mgrid@ncols * 4] # minish the new grid

ids_name <- mgridn_subset[ID %in% ids]
ids_name[, (c("ID", "x", "y", "fwts", "avail")) := NULL]

## Write the names into database
# user <- params$mapper$db_username
# password <- params$mapper$db_password

con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(),
                      dbname = params$mapper$db_sandbox_name,   
                      user = user,
                      password = password)

dbWriteTable(con, "oldkml", ids_name, overwrite = TRUE)
```
