---
title: "score_top_maps"
author: "Boka Luo"
date: "December 1, 2020"
output: 
  html_document:
    margin: auto
editor_options: 
  chunk_output_type:view_in_new_windows
---

## Packages
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(rmapaccuracy)
library(dplyr)
library(sf)
library(leaflet)
library(DBI)
source(here::here('spatial/R/tools/display_functions.R'))
```

## Arguments
```{r}
review.max <- 2 # maximum review time for each map
read_dbase <- TRUE

params <- yaml::yaml.load_file(here::here("common/config.yaml"))
dinfo <- params$labeller

reviewer.email <- dinfo$db_reviewer_email
host <- dinfo$db_host
con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(), host = host, 
                      dbname = "Africa", user = dinfo$db_username, 
                      password = dinfo$db_password)

reviewer.id <- dbGetQuery(
  con,
  paste0("SELECT id FROM users WHERE email='", reviewer.email, "'")
) %>% pull(id)
# reviewer_id <- tbl(con, "users") %>% filter(email == reviewer.email) %>% 
#   pull(id)
```


## Finished grids and top assignments
```{r, eval = read_dbase, echo = FALSE}
# Grids mapped by all 4 mappers
names.finish <- dbGetQuery(
  con,
  "SELECT name FROM kml_data WHERE mapped_count=4 AND kml_type='F'"
)

# Mapper scores from accuracy_data
assess <- dbGetQuery(
  con, 
  paste0("SELECT worker_id, new_score, count_acc, fragmentation_acc,",
         " edge_acc, ingrid_acc, outgrid_acc, field_skill, nofield_skill", 
         " FROM accuracy_data INNER JOIN assignment_data USING", 
         " (assignment_id)")
) %>% group_by(worker_id) %>% 
  summarise_all('mean') %>% 
  mutate(edge_score = 0.5 * fragmentation_acc + 0.5 * edge_acc)

#Associated approved assignments pending for review that is not reviewed by the reviewer
assignments <- dbGetQuery(
  con,
  paste0("SELECT name, hit_id, assignment_id, worker_id, delete_time,", 
         " reviewed_count FROM hit_data INNER JOIN assignment_data", 
         " USING (hit_id) WHERE name IN ('",
         paste(names.finish$name, collapse = "','"), "') ",
         "AND status='Approved' ",
         "AND reviewed_count <=", review.max, " ",
         "AND assignment_id NOT IN (SELECT assignment_id FROM",
         " assignment_review WHERE reviewer_id=", reviewer.id,")")
) %>% 
  left_join(
    .,
    assess %>% select(worker_id, new_score, edge_score), 
    by=c("worker_id")
  )

# Top assignments
assignments.top.score <- assignments %>% 
  group_by(hit_id) %>% 
  rename(score = new_score) %>% 
  filter(score == max(score)) %>% 
  select(-edge_score) %>% 
  ungroup()
  
assignments.top.edge <- assignments %>% 
  group_by(hit_id) %>% 
  filter(edge_score==max(edge_score)) %>% 
  select(-new_score) %>% 
  ungroup()

categories <- tbl(con, "categories") %>% collect()

# names.grouped.edge <- assignments.top.edge %>% filter(edge_group != "") %>% pull(name)
# names.grouped.binary <- assignments.top.score %>% filter(edge_group != "") %>% pull(name)

names.reviewed <- c()

```



### Review starts here
## get one uncommented map 

```{r}
name.select <- names.finish %>% 
  filter(name %in% assignments$name & !name %in% names.reviewed) %>% 
  sample_n(1) %>% 
  pull(name)

# name.select <- "TG0093561"
# Grid
grid_sel <- dbGetQuery(
  con,
  paste0("SELECT * FROM master_grid WHERE name='", name.select[1], "'")
) %>% data.table::data.table(.)
gcs <- "+proj=longlat +datum=WGS84 +no_defs"
gpoly <- point_to_gridpoly(grid_sel, w = 0.005 / 2, gcs, gcs)

# Scenes
scene <- dbGetQuery(
  con,
  paste0("SELECT season, tms_url FROM scenes_data WHERE cell_id=", grid_sel$id)
)
os_url <- scene %>% filter(season=="OS") %>% pull(tms_url)
gs_url <- scene %>% filter(season=="GS") %>% pull(tms_url)
```

## edge mode
### Check the map that has highest edge_acc and fragmentation_acc
```{r paged.print=TRUE}
options(viewer = NULL)
assignment.select.edge <- assignments.top.edge %>% filter(name==name.select)
aid.edge <- assignment.select.edge %>% pull(assignment_id)

# pick up fields for assignment
sqls <- paste0("select name, category, geom_clean",
               " FROM user_maps INNER JOIN categories ",
               "USING (category) where assignment_id='",
               aid.edge, "' ",
               "AND categ_group ='field'")
fields <- dbGetQuery(con, gsub(", geom_clean", "", sqls))
if(nrow(fields) > 0) {
  user_polys <- st_read(con, query = sqls)
} else {
  user_polys <- NULL
}

display_imagery(gpoly, user_polys, qaqc_maps=NULL, os_url, gs_url, categories)

```


## Feedback
### Binary
#### must be in range of [1,5], where correctly classified fields
. 1 -- lower than 65%
. 2 -- between 65% to 80%
. 3 -- over 80%; or the grid has field but the missing is small enough to be ignored
. 4 -- grid only has negative class and is 100% correct
. 5 -- lower than 65% but contains land cover type lacking for modelling, including dense fields, bare land, fallows, or cities
### Edge
#### error count is limited to range [0,10], if the error counts over 10, then annotate it as NULL
. 1 or more vertex missing of 1 field counts to 1 error
. 1 field missing counts to 1 error
. 1 edge missing between 2 fields counts to 1 error
#### error ratio is limited to [0, 1,0]; if the error ratio exceed this range, annotate it as NULL
. calculated by (edge error / total true fields), where
. 2 fields dissolved into 1 only count 1 on numerator
#### comment is required when both edge.errcount or edge.errratio are null, 
#### or it is all correct negative class
#### could be
. "'suitable'" -- edge errors less than 3 & less than 20% of true edges or is only a very tiny part of target grid
. "'suitable for fixing'" --edge erros less than 6, and is not hard to fix
. "'suitable but all negative'" -- the grid has no error but 100% negative
. "'not suitable'" -- edge errors over 6, and is hard to fix
. "'not sure'" -- edges are "not suitable" but contains land cover types lacking for modeling, including dense fields, bare land, fallows, water, or cities
```{r, echo=FALSE}
feedback.binary <- 4
feedback.edge.errcount <- 0 # numeric or "NULL"
feedback.edge.errratio <- 0.0 # numeric or "NULL"
feedback.comment <- "'suitable but all negative'"
```


### write the feedback to database
```{r, echo=FALSE}

dbExecute(
  con,
  sprintf(
    paste0("insert into assignment_review",
           "(reviewer_id, review_time, binary_group, edge_error_count,", 
           "edge_error_ratio, edge_comment, assignment_id) values",
           "(%d, '%s', %s, %s, %s, %s, %d);"),
    reviewer.id,
    as.character(Sys.time()),
    feedback.binary,
    feedback.edge.errcount,
    feedback.edge.errratio,
    feedback.comment,
    aid.edge
  )
)

# update reviewed_count in assignment_data
dbExecute(con,
          paste0("UPDATE assignment_data ",
                 "SET reviewed_count=",
          "(SELECT COUNT(1) FROM assignment_review ",
          "WHERE assignment_id=", aid.edge,")",
          "WHERE assignment_id=", aid.edge))

names.reviewed <- append(names.reviewed, assignment.select.edge$name) %>% 
  unique()
```


## binary mode
### See if the map has the highest score is different from that one in edge mode
### if yes, run following chunks for binary mode
### else, skip the binary mode
```{r}

assignment.select.binary <- assignments.top.score %>% 
  filter(name == name.select)
aid.binary <- assignment.select.binary %>% pull(assignment_id)

if (aid.binary == aid.edge) {
  print(paste0("Binary mode and edge mode are pointing to the same assignment.", 
               " Please skip following chunks"))
  names.reviewed <- append(names.reviewed, 
                           assignment.select.edge$name) %>% unique()
  
} else {
  print(paste0("Binary mode and edge mode are pointing to different", 
               " assignments. Please run following chunks"))
  
}
```

### Display imagery
```{r}
# pick up fields for assignment
sqls <- paste0("select name, category, geom_clean",
               " FROM user_maps INNER JOIN categories ",
               "USING (category) where assignment_id='",
               aid.binary, "' ",
               "AND categ_group ='field'")
fields <- dbGetQuery(con, gsub(", geom_clean", "", sqls))
if(nrow(fields) > 0) {
  user_polys <- st_read(con, query = sqls)
} else {
  user_polys <- NULL
}

display_imagery(gpoly, user_polys, qaqc_maps=NULL, os_url, gs_url, categories)

```


## Feedback
### Binary
#### must be in range of [1,5], where correctly classified fields
. 1 -- lower than 65%
. 2 -- between 65% to 80%
. 3 -- over 80%; or the grid has field but the missing is small enough to be ignored
. 4 -- grid only have negative class and is 100% correct
. 5 -- lower than 65% but contains land cover type lacking for modelding, including dense fields, bare land, fallows, or cities
### Edge
#### error count is limited to range [0,10], if the error counts over 10, then annotate it as NULL
. 1 or more vertex missing of 1 field counts to 1 error
. 1 field missing counts to 1 error
. 1 edge missing between 2 fields counts to 1 error
#### error ratio is limited to [0, 1,0]; if the error ratio exceed this range, annotate it as NULL
. calculated by (edge error / total true fields), where
. 2 fields dissolved into 1 only count 1 on numerator
#### comment is required when both edge.errcount or edge.errratio are null, 
#### or it is all correct negative class
#### could be
. "'suitable'" -- edge errors less than 3 & less than 20% of true edges or is only a very tiny part of target grid
. "'suitable for fixing'" --edge erros less than 6, and is not hard to fix
. "'suitable but all negative'" -- the grid has no error but 100% negative
. "'not suitable'" -- edge errors over 6, and is hard to fix
. "'not sure'" -- edges are "not suitable" but contains land cover types lacking for modeling, including dense fields, bare land, fallows, water, or cities
```{r, echo=FALSE}
feedback.binary <- 4
feedback.edge.errcount <- 0
feedback.edge.errratio <- 0.0 # numeric or "NULL"
feedback.comment <- "'suitable but all negative'"
```


### write feedback to database
```{r, echo=FALSE}

dbExecute(
  con,
  sprintf(
    paste0("insert into assignment_review",
           "(reviewer_id, review_time, binary_group,", 
           " edge_error_count, edge_error_ratio, edge_comment,", 
           " assignment_id) values",
           " (%d, '%s', %s, %s, %s, %s, %d);"),
    reviewer.id,
    as.character(Sys.time()),
    feedback.binary,
    feedback.edge.errcount,
    feedback.edge.errratio,
    feedback.comment,
    aid.binary
  )
)

# update reviewed_count in assignment_data
dbExecute(
  con,
  paste0("UPDATE assignment_data",
         " SET reviewed_count=",
         "(SELECT COUNT(1) FROM assignment_review",
         " WHERE assignment_id=", aid.binary, ")",
         "WHERE assignment_id=", aid.binary)
)

names.reviewed <- append(names.reviewed, assignment.select.binary$name) %>% unique()
```

## close connection
### Run this chunk when stop assessment
```{r, echo=FALSE}
dbDisconnect(con)
```